### 2023-01-30 Strydom_et_al_2023_graph_embedding_draft.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@3032c689da6f52ef691456e9389c596b3e43be49 🚀

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Graph embedding and transfer learning can help
predict species interaction networks despite data limitations}
\vskip 5em

@@

\begin{enumerate}
    \item Metawebs, (networks of potential interactions within a species
pool) are a powerful abstraction to understand how large-scale species
interaction networks are structured.%
    \item Because metawebs are typically expressed at large spatial and
taxonomic scales, assembling them is a tedious and costly process;
predictive methods can help circumvent the limitations in data
deficiencies, by providing `draft' metawebs.%
    \item One way to improve our ability to predict metawebs is to
maximize available information by using graph embeddings, instead of the
list of species interactions. Graph embedding is an emerging field in
machine learning that holds great potential for ecological problems.%
    \item In this perspective, we outline how the challenges associated
with inferring metawebs line-up with the advantages of graph embeddings;
as well as discuss how the choice of the species pool has consequences
on the reconstructed network, but also embeds hypotheses about which
human-made boundaries are ecologically meaningful.%
\end{enumerate}

@@

The ability to infer \emph{potential} interactions could serve as a
significant breakthrough in our ability to conceptualize species
interaction networks over large spatial scales (Hortal et al., 2015).
Robust inferences would not only boost our understanding of the
structure of species interaction networks, but also increase the amount
of information that can be used for biodiversity management. In a recent
overview of the field of ecological network prediction, Strydom,
Catchen, et al. (2021) identified two challenges of interest to the
prediction of interactions at large scales. First, there is a relative
scarcity of relevant data in most places globally -- paradoxically, this
restricts our ability to infer interactions to locations where inference
is perhaps the least required (and leaves us unable to make inference in
data scarce regions); second, accurate predictors are important for
accurate predictions, and the lack of methods that can leverage a small
amount of \emph{accurate} data is a serious impediment to our predictive
ability. In most places, our most reliable biodiversity knowledge is
that of a species pool (\emph{i.e.} a set of potentially interacting
species in a given area): through the analysis of databases like the
Global Biodiversity Information Facility (GBIF) or the International
Union for the Conservation of Nature (IUCN), it is possible to construct
a list of species for a region of interest; however inferring the
potential interactions between these species still remains a challenge.

Following the definition of Dunne (2006), a metaweb is the ecological
network analogue to the species pool; specifically, it inventories all
\emph{potential} interactions between species for a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb itself is not a prediction of the network at a specific point
within the spatial area it covers: it will have a different structure,
notably by having a larger connectance (see \emph{e.g.} Wood et al.,
2015) and complexity (see \emph{e.g.} Galiana et al., 2022), from any of
these local networks. These local networks (which capture the \(\alpha\)
diversity of interactions) are a subset of the metaweb's species and
realized interactions, and have been called ``metaweb realizations''
(Poisot et al., 2015). Differences between local networks and their
metawebs are due to chance, species abundance and co-occurrence, local
environmental conditions, and local distribution of functional traits,
among others. Specifically, although co-occurrence can be driven by
interactions (Cazelles et al., 2016), co-occurrence alone is not a
predictor of interactions (Blanchet et al., 2020; Thurman et al., 2019),
and therefore lack of co-occurrence cannot be used to rule out lack of a
feasible interaction. Yet, recent results by Saravia et al. (2021)
strongly suggest that the local (metaweb) realizations only respond
weakly to local conditions: instead, they reflect constraints inherited
by the structure of their metaweb. This establishes the metaweb
structure as the core goal of predictive network ecology, as it is a
required information to accurately produce downscaled, local
predictions.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it
represents the ``upper bounds'' on what the composition of the local
networks, given a local species pool, can be (see \emph{e.g.} McLeod et
al., 2021); this information can help evaluate the ability of ecological
assemblages to withstand the effects of, for example, climate change
(Fricke et al., 2022). These local networks may be reconstructed given
an appropriate knowledge of local species composition and provide
information on the structure of food webs at finer spatial scales. This
has been done for example for tree-galler-parasitoid systems (Gravel et
al., 2018), fish trophic interactions (Albouy et al., 2019), tetrapod
trophic interactions (J. Braga et al., 2019; O'Connor et al., 2020), and
crop-pest networks (Grünig et al., 2020). In this contribution, we
highlight the power of viewing (and constructing) metawebs as
\emph{probabilistic} objects in the context of rare interactions,
discuss how a family of machine learning tools (graph embeddings and
transfer learning) can be used to overcome data limitations to metaweb
inference, and highlight how the use of metawebs introduces important
questions for the field of network ecology.

@@

Treating interactions as probabilistic (as opposed to binary) events is
a more nuanced and realistic way to represent them. Dallas et al. (2017)
suggested that most links in ecological networks are cryptic,
\emph{i.e.} uncommon or hard to observe. This argument echoes Jordano
(2016): sampling ecological interactions is difficult because it
requires first the joint observation of two species, and then the
observation of their interaction. In addition, it is generally expected
that weak or rare links will be more prevalent in networks than common
or strong links (Csermely, 2004), compared to strong, persistent links;
this is notably the case in food chains, wherein many weaker links are
key to the stability of a system (Neutel et al., 2002). In the light of
these observations, we expect to see an over-representation of
low-probability (rare) interactions under a model that accurately
predicts interaction probabilities.

@@

Graph (or Network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Arsov \& Mirceva,
2019; Chami et al., 2022). Their main goal is to learn a low dimensional
vector representations of the graph (embeddings), such that its key
properties (\emph{e.g.} local or global structures) are retained in the
embedding space (Yan et al., 2005). The embedding space may, but will
not necessarily, have lower dimensionality than the graph. Ecological
networks are promising candidates for the routine application of
embeddings, as they tend to possess a shared structural backbone (see
\emph{e.g.} Bramon Mora et al., 2018), which hints at structural
invariants in empirical data. Assuming that these structural invariants
are common enough, they would dominate the structure of networks, and
therefore be adequately captured by the first (lower) dimensions of an
embedding, without the need to measure derived aspects of their
structure (\emph{e.g.} motifs, paths, modularity, \ldots).

@@

Before moving further, it is important to clarify the epistemic status
of node values derived from embeddings: specifically, they are
\emph{not} functional traits, and therefore should not be discussed in
terms of effects or responses. As per the framework of Malaterre et al.
(2019), these values neither derive from, nor result in, changes in
organismal performance, and should therefore not be used to quantify
\emph{e.g.} functional diversity. This holds true even when there are
correlations between latent values and functional traits: although these
enable an ecological discussion of how traits condition the structure of
the network, the existence of a statistical relationship does not
elevate the latent values to the status of functional traits.

@@

Indeed, food webs are inherently low-dimensional objects, and can be
adequately represented with less than ten dimensions (J. Braga et al.,
2019; M. P. Braga et al., 2021; Eklöf et al., 2013). Simulation results
by Botella et al. (2022) suggests that there is no dominant method to
identify architectural similarities between networks: multiple
approaches need to be tested and compared to the network descriptor of
interest on a problem-specific basis. This matches previous results on
graph embedding, wherein different embedding algorithms yield different
network embeddings (Goyal \& Ferrara, 2018), calling for a careful
selection of the problem-specific approach to use. In
tbl.~\ref{tbl:methods}, we present a selection of common graph and node
embedding methods, alongside examples of their use to predict
interactions or statistical associations between species. These methods
rely largely on linear algebra or pseudo-random walks on graphs. All
forms of embeddings presented in the table share the common property of
summarizing their objects into (sets of) dense feature vectors, that
capture the overall network structure, pairwise information on nodes,
and emergent aspects of the network, in a compressed way (\emph{i.e.}
with some information loss, as we later discuss in the illustration).
Node embeddings tend to focus on maintaining pairwise relationships
(\emph{i.e.} species interactions), while graph embeddings focus on
maintaining the network structure (\emph{i.e.} emergent properties).
Nevertheless, some graph embedding techniques (like RDPG, see
\emph{e.g.} Wu et al., 2021) will provide high-quality node-level
embeddings while also preserving network structure.

@@

The results of fig.~\ref{fig:illustration1} show that we can extract an
embedding of the metaweb that captures enough variance to be relevant;
specifically, this is true both of \(L_2\) loss (indicating that RDPG is
able to capture pairwise processes) and the cumulative variance
explained (indicating that RDPG is able to capture network-level
structure). Therefore, in fig.~\ref{fig:illustration2}, we relate the
values of latent variables for hosts to different ecologically-relevant
data. In panel \textbf{A}, we show that host with a higher value on the
first dimension have fewer parasites. This relates to the body size of
hosts in the \emph{PanTHERIA} database (Jones et al., 2009), as shown in
panel \textbf{B}: interestingly, the position on the first axis is only
weakly correlated to body mass of the host; this matches well
established results showing that body size/mass is not always a direct
predictor of parasite richness in terrestrial mammals (Morand \& Poulin,
1998), a result we observe in panel \textbf{C}. Finally, in panel
\textbf{D}, we can see how different taxonomic families occupy different
positions on the first axis, with \emph{e.g.} Sciuridae being biased
towards higher values. These results show how we can look for ecological
informations in the output of network embeddings, which can further be
refined into the selection of predictors for transfer learning.

@@

If the initial metaweb is too narrow in scope, notably from a taxonomic
point of view, the chances of finding another area with enough related
species (through phylogenetic relatedness or similarity of functional
traits) to make a reliable inference decreases. This is because transfer
requires similarity (fig.~\ref{fig:embedding}). A diagnostic for the
lack of similar species would likely be large confidence intervals
during estimation of the values in the low-rank space. In other words,
the representation of the original graph is difficult to transfer to the
new problem. Alternatively, if the initial metaweb is too large
(taxonomically), then the resulting embeddings would need to represent
interactions between taxonomic groups that are not present in the new
location. This would lead to a much higher variance in the starting
dataset, and to under-dispersion in the target dataset, resulting in the
potential under or over estimation of the strength of new predicted
interactions. Llewelyn et al. (2022) provide compelling evidence for
these situations by showing that, even at small spatial scales, the
transfer of information about interactions becomes more challenging when
areas rich with endemic species are considered. The lack of well
documented metawebs is currently preventing the development of more
concrete guidelines. The question of phylogenetic relatedness and
distribution is notably relevant if the metaweb is assembled in an area
with mostly endemic species (\emph{e.g.} a system that has undergone
recent radiation or that has remained in isolation for a long period of
time might not have an analogous system with which to draw knowledge
from), and as with every predictive algorithm, there is room for the
application of our best ecological judgement. Because this problem
relates to distribution of species in the geographic or phylogenetic
space, it can certainly be approached through assessing the performance
of embedding transfer in simulated starting/target species pools.

@@

The area for which we seek to predict the metaweb should determine the
species pool on which the embedding is performed. Metawebs can be
constructed by assigning interactions in a list of species within
geographic boundaries. The upside of this approach is that information
relevant for the construction of this dataset is likely to exist, as
countries usually set conservation goals at the national level (Buxton
et al., 2021), and as quantitative instruments are consequently designed
to work at these scales (Turak et al., 2017); specific strategies are
often enacted at smaller scales, nested within a specific country (Ray
et al., 2021). But there is no guarantee that these boundaries are
meaningful. In fact, we do not have a satisfying answer to the question
of ``where does an ecological network stop?'', the answer to which would
dictate the spatial span to embed/predict. Recent results by Martins et
al. (2022) suggest that networks are shaped within eco-regions, with
abrupt structural transitions from an eco-region to the next. Should
this trend hold generally, this would provide an ecologically-relevant
scale at which metawebs can be downscaled and predicted. Other solutions
could leverage network-area relationships to identify areas in which
networks are structurally similar (see \emph{e.g.} Fortin et al., 2021;
Galiana et al., 2022, 2018). Both of these solutions requires ample
pre-existing information about the network in space. Nevertheless, the
inclusion of species for which we have data but that are not in the
right spatial extent\emph{may} improve the performance of approaches
based on embedding and transfer, \emph{if} they increase the similarity
between the target and destination network. This proposal can
specifically be evaluated by adding nodes to the network to embed, and
evaluating the performance of predictive models (see \emph{e.g.}
Llewelyn et al., 2022).

@@

In large parts of the world, geographic boundaries only reflect the
legacy of settler colonialism, which drives global disparity in capacity
to collect and publish ecological data. Applying any embedding to biased
data does not debias them, but rather embeds these biases, propagating
them to the models using embeddings to make predictions. Furthermore,
the use of ecological data itself is not an apolitical act (Nost \&
Goldstein, 2021): data infrastructures tend to be designed to answer
questions within national boundaries (therefore placing contingencies on
what is available to be embedded), their use often drawing upon, and
reinforcing, territorial statecraft (see \emph{e.g.} Barrett, 2005). As
per Machen \& Nost (2021), these biases are particularly important to
consider when knowledge generated algorithmically is used to supplement
or replace human decision-making, especially for governance (\emph{e.g.}
enacting conservation decisions on the basis of model prediction). As
information on networks is increasingly leveraged for conservation
actions (see \emph{e.g.} Eero et al., 2021; Naman et al., 2022; Stier et
al., 2017), the need to appraise and correct biases that are unwittingly
propagated to algorithms when embedded from the original data is
immense. These considerations are even more urgent in the specific
context of biodiversity data. Long-term colonial legacies still shape
taxonomic composition to this day (Lenzner et al., 2022; Raja, 2022),
and much shorter-term changes in taxonomic and genetic richness of
wildlife emerged through environmental racism (Schmidt \& Garroway,
2022). These considerations mean that ``the set of species found at a
place'' is not only a response to ecological processes, and therefore
there may exist forcing on the interactions between these species.

>>>>

\fancypagestyle{normal}
{
  \fancyhf{}
  \fancyfoot[R]{\footnotesize\sffamily\thepage\ of \pageref*{LastPage}}
}
\begin{document}
\raggedright
\thispagestyle{empty}
{\Large\bfseries\sffamily Graph embedding and transfer learning can help
predict potential species interaction networks despite data limitations}
\vskip 5em

@@

\begin{enumerate}
    \item Metawebs, (networks of potential interactions within a species
pool) are a powerful abstraction to understand how large-scale species
interaction networks are structured.%
    \item Because metawebs are typically expressed at large spatial and
taxonomic scales, assembling them is a tedious and costly process;
predictive methods can help circumvent the limitations in data
deficiencies, by providing a `first approximation' of metawebs.%
    \item One way to improve our ability to predict metawebs is to
maximize available information by using graph embeddings, as opposed to
an exhaustive list of species interactions. Graph embedding is an
emerging field in machine learning that holds great potential for
ecological problems.%
    \item Here, we outline how the challenges associated with inferring
metawebs line-up with the advantages of graph embeddings; followed by a
discussion as to how the choice of the species pool has consequences on
the reconstructed network, specifically as to the role of human-made (or
arbitrarily assigned) boundaries and how these my influence ecological
hypotheses.%
\end{enumerate}

@@

The ability to infer \emph{potential} interactions could serve as a
significant breakthrough in our ability to conceptualize species
interaction networks over large spatial scales (Hortal et al., 2015).
Reliable inferences would not only boost our understanding of the
structure of species interaction networks, but also increase the amount
of information that can be used for biodiversity management. In a recent
overview of the field of ecological network prediction, Strydom,
Catchen, et al. (2021) identified two challenges of interest to the
prediction of interactions at large scales. First, there is a relative
scarcity of relevant data in most places globally -- restricting the
inference of interactions to locations where least required (and leaves
us unable to make inference in data scarce regions); second, accurate
predictors are important for accurate predictions, and the lack of
methods that can leverage a small amount of \emph{accurate} data is a
serious impediment to our predictive ability. In most places, our most
reliable biodiversity knowledge is that of a species pool where a set of
potentially interacting species in a given area could occur: through the
analysis of databases like the Global Biodiversity Information Facility
(GBIF) or the International Union for the Conservation of Nature (IUCN),
it is possible to construct a list of species for a region of interest;
however inferring the potential interactions between these species still
remains a challenge.

Following the definition of Dunne (2006), a metaweb is the ecological
network analogue to the species pool; specifically, it inventories all
\emph{potential} interactions between species for a spatially delimited
area (and so captures the \(\gamma\) diversity of interactions). The
metaweb itself is not a prediction of local networks at specific
locations within the spatial area it covers: it will have a different
structure, notably by having a larger connectance (see \emph{e.g.} Wood
et al., 2015) and complexity (see \emph{e.g.} Galiana et al., 2022),
from any of these local networks. These local networks (which capture
the \(\alpha\) diversity of interactions) are a subset of the metaweb's
species and realized interactions, and have been called ``metaweb
realizations'' (Poisot et al., 2015). Differences between local networks
and their metawebs are due to chance, species abundance and
co-occurrence, local environmental conditions, and local distribution of
functional traits, among others. Specifically, although co-occurrence
can be driven by interactions (Cazelles et al., 2016), co-occurrence
alone is not a predictor of interactions (Blanchet et al., 2020; Thurman
et al., 2019), and therefore lack of co-occurrence cannot be used to
rule out lack of a feasible interaction. Yet, recent results by Saravia
et al. (2021) strongly suggested that the local (metaweb) realizations
only respond weakly to local conditions: instead, they reflect
constraints inherited by the structure of their metaweb. This sets up
the core goal of predictive network ecology as the prediction of metaweb
structure, as it is required to accurately produce downscaled, local
predictions.

Because the metaweb represents the joint effect of functional,
phylogenetic, and macroecological processes (Morales-Castilla et al.,
2015), it holds valuable ecological information. Specifically, it
represents the ``upper bounds'' on what the composition of the local
networks, given a local species pool, can be (see \emph{e.g.} McLeod et
al., 2021); this information can help evaluate the ability of ecological
assemblages to withstand the effects of, for example, climate change
(Fricke et al., 2022). These local networks may be reconstructed given
an appropriate knowledge of local species composition and provide
information on the structure of food webs at finer spatial scales. This
has been done for example for tree-galler-parasitoid systems (Gravel et
al., 2018), fish trophic interactions (Albouy et al., 2019), tetrapod
trophic interactions (J. Braga et al., 2019; O'Connor et al., 2020), and
crop-pest networks (Grünig et al., 2020). In this contribution, we
highlight the power of viewing (and constructing) metawebs as
\emph{probabilistic} objects in the context of low-probability
interactions, discuss how a family of machine learning tools (graph
embeddings and transfer learning) can be used to overcome data
limitations to metaweb inference, and highlight how the use of metawebs
introduces important questions for the field of network ecology.

@@

Treating interactions as probabilistic (as opposed to binary) events is
a more nuanced and realistic way to represent them. Dallas et al. (2017)
suggested that most links in ecological networks are cryptic,
\emph{i.e.} uncommon or hard to observe. This argument echoes Jordano
(2016): sampling ecological interactions is difficult because it
requires first the joint observation of two species, and then the
observation of their interaction. In addition, it is generally expected
that weak or rare links will be more prevalent in networks than common
or strong links (Csermely, 2004), compared to strong, persistent links;
this is notably the case in food chains, wherein many weaker links are
key to the stability of a system (Neutel et al., 2002). In the light of
these observations, we expect to see an over-representation of
low-probability (hereafter rare) interactions under a model that
accurately predicts interaction probabilities.

@@

Graph (or network) embedding (fig.~\ref{fig:embedding}) is a family of
machine learning techniques, whose main task is to learn a mapping
function from a discrete graph to a continuous domain (Arsov \& Mirceva,
2019; Chami et al., 2022). Their main goal is to learn a low dimensional
vector representations of the graph (embeddings), such that its key
properties (\emph{e.g.} local or global structures) are retained in the
embedding space (Yan et al., 2005). The embedding space may, but will
not necessarily, have lower dimensionality than the graph. Ecological
networks are promising candidates for the routine application of
embeddings, as they tend to possess a shared structural backbone (see
\emph{e.g.} Bramon Mora et al., 2018), which hints at structural
invariants in empirical data. Assuming that these structural invariants
are common enough, they would dominate the structure of networks, and
therefore be adequately captured by the first (lower) dimensions of an
embedding, without the need to measure derived aspects of their
structure (\emph{e.g.} motifs, paths, modularity, \ldots).

@@

Before moving further, it is important to clarify the epistemic status
of node values derived from embeddings: specifically, they are
\emph{not} functional traits, and therefore should not be interpreted in
terms of effects or responses. As per the framework of Malaterre et al.
(2019), these values neither derive from, nor result in, changes in
organismal performance, and should therefore not be used to quantify
\emph{e.g.} functional diversity. This holds true even when there are
correlations between latent values and functional traits: although these
enable an ecological discussion of how traits condition the structure of
the network, the existence of a statistical relationship does not
elevate the latent values to the status of functional traits.

@@

Food webs are inherently low-dimensional objects, and can be adequately
represented with less than ten dimensions (J. Braga et al., 2019; M. P.
Braga et al., 2021; Eklöf et al., 2013). Simulation results by Botella
et al. (2022) suggested that there is no dominant method to identify
architectural similarities between networks: multiple approaches need to
be tested and compared to the network descriptor of interest on a
problem-specific basis. This matches previous results on graph
embedding, wherein different embedding algorithms yield different
network embeddings (Goyal \& Ferrara, 2018), calling for a careful
selection of the problem-specific approach to use. In
tbl.~\ref{tbl:methods}, we present a selection of common graph and node
embedding methods, alongside examples of their use to predict
interactions or statistical associations between species. These methods
rely largely on linear algebra or pseudo-random walks on graphs. All
forms of embeddings presented in tbl.~\ref{tbl:methods} share the common
property of summarizing their objects into (sets of) dense feature
vectors, that capture the overall network structure, pairwise
information on nodes, and emergent aspects of the network, in a
compressed way (\emph{i.e.} with some information loss, as we later
discuss in the illustration). Node embeddings tend to focus on
maintaining pairwise relationships (\emph{i.e.} species interactions),
while graph embeddings focus on maintaining the network structure
(\emph{i.e.} emergent properties). Nevertheless, some graph embedding
techniques (like RDPG, see \emph{e.g.} Wu et al., 2021) will provide
high-quality node-level embeddings while also preserving network
structure.

@@

The results of fig.~\ref{fig:illustration1} show that we can extract an
embedding of the metaweb that captures enough variance to be relevant;
specifically, this is true for both \(L_2\) loss (indicating that RDPG
is able to capture pairwise processes) and the cumulative variance
explained (indicating that RDPG is able to capture network-level
structure). Therefore, in fig.~\ref{fig:illustration2}, we relate the
values of latent variables for hosts to different ecologically-relevant
data. In panel \textbf{A}, we show that host with a higher value on the
first dimension have fewer parasites. This relates to the body size of
hosts in the \emph{PanTHERIA} database (Jones et al., 2009), as shown in
panel \textbf{B}: interestingly, the position on the first axis is only
weakly correlated to body mass of the host; this matches well
established results showing that body size/mass is not always a direct
predictor of parasite richness in terrestrial mammals (Morand \& Poulin,
1998), a result we observe in panel \textbf{C}. Finally, in panel
\textbf{D}, we can see how different taxonomic families occupy different
positions on the first axis, with \emph{e.g.} Sciuridae being biased
towards higher values. These results show how we can look for ecological
informations in the output of network embeddings, which can further be
refined into the selection of predictors for transfer learning.

@@

If the initial metaweb is too narrow in scope, notably from a taxonomic
point of view, the chances of finding another area with enough related
species (through phylogenetic relatedness or similarity of functional
traits) to make a reliable inference decreases. This is because transfer
requires similarity (fig.~\ref{fig:embedding}). A diagnostic for the
lack of similar species would likely be large confidence intervals
during estimation of the values in the low-rank space. In other words,
the representation of the original graph is difficult to transfer to the
new problem. Alternatively, if the initial metaweb is too large
(taxonomically), then the resulting embeddings would need to represent
interactions between taxonomic groups that are not present in the new
location. This would lead to a much higher variance in the starting
dataset, and to under-dispersion in the target dataset, resulting in the
potential under or over estimation of the strength of new predicted
interactions. Llewelyn et al. (2022) provided compelling evidence for
these situations by showing that, even at small spatial scales, the
transfer of information about interactions becomes more challenging when
areas rich with endemic species are considered. The lack of well
documented metawebs is currently preventing the development of more
concrete guidelines. The question of phylogenetic relatedness and
distribution is notably relevant if the metaweb is assembled in an area
with mostly endemic species (\emph{e.g.} a system that has undergone
recent radiation or that has remained in isolation for a long period of
time might not have an analogous system with which to draw knowledge
from), and as with every predictive algorithm, there is room for the
application of our best ecological judgement. Because this problem
relates to distribution of species in the geographic or phylogenetic
space, it can certainly be approached through assessing the performance
of embedding transfer in simulated starting/target species pools.

@@

The area for which we seek to predict the metaweb should determine the
species pool on which the embedding is performed. Metawebs can be
constructed by assigning interactions in a list of species within
specific regions. The upside of this approach is that information
relevant for the construction of this dataset is likely to exist, as
countries usually set conservation goals at the national level (Buxton
et al., 2021), and as quantitative instruments are consequently designed
to work at these scales (Turak et al., 2017); specific strategies are
often enacted at smaller scales, nested within a specific country (Ray
et al., 2021). But there is no guarantee that these arbitrary boundaries
are meaningful. In fact, we do not have a satisfying answer to the
question of ``where does an ecological network stop?'', the answer to
which would dictate the spatial span to embed/predict. Recent results by
Martins et al. (2022) suggested that networks are shaped within
eco-regions, with abrupt structural transitions from an eco-region to
the next. Should this trend hold generally, this would provide an
ecologically-relevant scale at which metawebs can be downscaled and
predicted. Other solutions could leverage network-area relationships to
identify areas in which networks are structurally similar (see
\emph{e.g.} Fortin et al., 2021; Galiana et al., 2022, 2018). Both of
these solutions requires ample pre-existing information about the
network in space. Nevertheless, the inclusion of species for which we
have data but that are not in the right spatial extent\emph{may} improve
the performance of approaches based on embedding and transfer, \emph{if}
they increase the similarity between the target and destination network.
This proposal can specifically be evaluated by adding nodes to the
network to embed, and evaluating the performance of predictive models
(see \emph{e.g.} Llewelyn et al., 2022).

@@

In large parts of the world, boundaries that delineate geographic
regions are merely a reflection the legacy of settler colonialism, which
drives global disparity in capacity to collect and publish ecological
data. Applying any embedding to biased data does not debias them, but
rather embeds these biases, propagating them to the models using
embeddings to make predictions. Furthermore, the use of ecological data
itself is not an apolitical act (Nost \& Goldstein, 2021): data
infrastructures tend to be designed to answer questions within national
boundaries (therefore placing contingencies on what is available to be
embedded), their use often drawing upon, and reinforcing, territorial
statecraft (see \emph{e.g.} Barrett, 2005). As per Machen \& Nost
(2021), these biases are particularly important to consider when
knowledge generated algorithmically is used to supplement or replace
human decision-making, especially for governance (\emph{e.g.} enacting
conservation decisions on the basis of model prediction). As information
on networks is increasingly leveraged for conservation actions (see
\emph{e.g.} Eero et al., 2021; Naman et al., 2022; Stier et al., 2017),
the need to appraise and correct biases that are unwittingly propagated
to algorithms when embedded from the original data is immense. These
considerations are even more urgent in the specific context of
biodiversity data. Long-term colonial legacies still shape taxonomic
composition to this day (Lenzner et al., 2022; Raja, 2022), and much
shorter-term changes in taxonomic and genetic richness of wildlife
emerged through environmental racism (Schmidt \& Garroway, 2022). These
considerations mean that ``the set of species found at a place'' is not
only a response to ecological processes, and therefore there may exist
forcing on the interactions between these species.

### 2023-01-30 Strydom_et_al_2023_graph_embedding_preprint.tex
Deploying to gh-pages from @ PoisotLab/ms_metaweb_perspectives@3032c689da6f52ef691456e9389c596b3e43be49 🚀

\noindent {\LARGE \textbf{\textsf{Graph embedding and transfer learning
can help predict species interaction networks despite data
limitations}}}

>>>>

\noindent {\LARGE \textbf{\textsf{Graph embedding and transfer learning
can help predict potential species interaction networks despite data
limitations}}}
